---
title: "An Analysis of COVID-19 Cases by Country"
output: html_document
---
Given the countries and the count of positive, hospitalised, recovered, deaths and other variables, this analysis aims to answer the question: Which countries have had the highest number of COVID-19 cases against the number of tests administered?
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Initial Data Exploration
```{r}
library(readr)
library(tidyverse)
library(tibble)
covid_df <- read_csv("covid19.csv")
```

###
```{r}
dim(covid_df)
```
```{r}
head(covid_df)
```
```{r}
vector_cols <- colnames(covid_df)
vector_cols
```
glimpse(covid_df)

### Data Wrangling
```{r}
covid_df_all_states <- covid_df %>% filter(Province_State == 'All States') %>% select(-Province_State)
covid_df_all_states
```

### Isolating needed columns
To avoid confusion with cumulative and daily data, I decided to focus on the daily statistics and selected relevant columns.

```{r}
covid_df_all_states_daily <- covid_df_all_states %>% select(Date, Country_Region, active, hospitalizedCurr, daily_tested, daily_positive)
head(covid_df_all_states_daily)
```
## Extracting the Top Ten countries in the number of tested cases
### Summarizing the data grouping by `Country_Region` column.
```{r}
covid_df_all_states_daily_sum <- covid_df_all_states_daily %>% group_by(Country_Region) %>% summarize(tested = sum(daily_tested),
positive = sum(daily_positive), active = sum(active), hospitalized = sum(hospitalizedCurr)) %>% arrange(-tested)
covid_df_all_states_daily_sum
```
### Extracting top 10 most number of tested cases.
```{r}
covid_top_10 <- head(covid_df_all_states_daily_sum, 10)
covid_top_10
```

## Identifying highest positive against tested cases
### Creating and naming vectors
```{r}
countries <- covid_top_10$Country_Region
tested_cases <- covid_top_10$tested
positive_cases <- covid_top_10$positive
active_cases <- covid_top_10$active
hospitalized_cases <- covid_top_10$hospitalized
```
```{r}
names(positive_cases) <- countries
names(tested_cases) <- countries
names(active_cases) <- countries
names(hospitalized_cases) <- countries
```

### Calculating top 3 positive cases
### Ratio of positive to total cases per country
```{r}
positive_cases
sum(positive_cases)
mean(positive_cases)
positive_cases / sum(positive_cases)
```
### Ratio of positive against total tested cases
```{r}
positive_cases / tested_cases
```

### Findings for top 3 positive cases
```{r}
positive_tested_top_3 <- c("United States" = 0.11, "United Kingdom" = 0.11, "Turkey" = 0.08)
positive_tested_top_3
```

## Keeping Relevant Information
```{r}
united_kingdom <- c(0.11, 1473672, 166909, 0, 0)
united_states <- c(0.10, 17282363, 1877179, 0, 0)
turkey <- c(0.08, 2031192, 163941, 2980960, 0)
```
### Creating COVID matrix
```{r}
covid_mat <- rbind(united_states, united_kingdom, turkey)
colnames(covid_mat) <- c("Ratio", "tested", "positive", "active", "hospitalized")
covid_mat
```

## Summary
```{r}
Question <- "Which countries have had the highest number of positive cases against the number of tests?"
Answer <- c("Positive tested cases" = positive_tested_top_3)
datasets <- list(original = covid_df, all_states = covid_df_all_states, daily_cases = covid_df_all_states_daily, top_10 = covid_top_10)

matrices <- list(covid_mat)
vectors <- list(vector_cols, countries)

```



